Algebraic filtering of surfaces from medical images with Julia

### 2019-11-11 src/paper.tex
paper.tex partially reviewed

@@

Some basic concepts of solid modeling, and in particular the foundational idea of representation scheme, as well as basic concepts of algebraic topology, are shortly introduced in this Section, including the computation of the matrix of a boundary operator between chain spaces.

@@

It is possible to show that the \textsc{lar} representation scheme is very expressive, i.e.~that it  has a large domain,  including collections of: line segments, quads, triangles, polygons, meshes;  pixels, voxels, volume images; B-reps, enumerative and decompositive representations of solids. 
In this paper we apply \textsc{lar} methods to computation of boundary representations of solid models from segmentation (labeling) of 3D medical images.
To display a triangulation of  boundary faces  in their proper position in space, the information required is contained in the \emph{geometric chain complex} (GCC):
\[
\mu: C_0\to\E^3,\ (\delta_0, \delta_1, \delta_2)
\qquad\equiv\qquad
\texttt{(geometry, topology) = (W, (EV, FE, CF))}
\]
The GCC allows to transform the (possibly non connected) boundary 2-cycle of surfaces as a standard B-rep~\cite{shapiroSM:202}. 
The \texttt{geometry} is given by
the embedding matrix \texttt{W} of vertices (0-cells), and \texttt{topology} is
given by the three sparse matrices \texttt{CF}, \texttt{FE}, \texttt{EV} of coboundaries
{i.e.},~$\delta_3, \delta_2, \delta_1$, of  chain complex describing a   
space arrangement~\cite{paoluzzi2019finite}.
Note that ordered pairs of letters from \texttt{V,E,F,C}, correspond to \emph{\emph{\texttt{V}}ertices$\to$\emph{\texttt{E}}dges$\to$\emph{\texttt{F}}aces$\to$\emph{\texttt{C}}ells} into the 
\emph{\emph{\texttt{C}}olumn$\to$\emph{\texttt{R}}ow} order of matrix maps of operators.

\subsubsection{Construction of boundary matrix $\partial_d$}

@@

\begin{figure}[tbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.75\linewidth]{figs/boundary_matrix_4x4x4.pdf} 
   \caption{A binary image of the coboundary operator  $\delta_2 = \partial_3^\top : C_2 \to C_3$, built for a small 3D image with shape $(4,4,4)$. Note that the number of rows equates the cardinality $4\times 4\times 4 = 64$ of the voxel set; the number of columns is $d\,n\,(1+n)^{d-1} = 3\times 4\times 25 = 300$. Of course, the number of non-zeros per row (cardinality of single voxel facets) is six, whereas the number of non-zeros per column is two, but on boundary facets.}
   \label{fig:boundary_matrix_4x4x4}
\end{figure}

@@

Every boundary chain extracted from an image block $\B(i,j,n)$ is a \emph{2-cycle}, i.e., a closed 2-chain---in other words, a 2-chain with empty boundary. Such 2-cycles are joined together by removing the double 2-cells (at the boundaries of bricks) after having suitably shifted their indices to an unique linear representation of the whole image. The resulting raster surface is made by mutually orthogonal raster facets, that must be smoothed in order to get a fair surface. A linear time and space algorithm for this purpose is the Laplacian smoothing, which iteratively  moves each vertex (0-cell) to the centroid of its neighbors. A well known weakness of this simple algorithm is the asymptotic convergence of the whole mesh to a single point, resulting in unfair size reduction even after few iterations.  The Taubin smoothing algorithm~\cite{Taubin:1995:SPA:218380.218473,egst.20001029} alternates two Laplacian smoothing steps with shrink and inflate effects respectively, with the result of letting pretty invariant sizes and volume of the smoothed mesh.
The best results are obtained on meshes which have small variations of edge length and face angles, like for surfaces extracted from 3D raster images, as in our case.

@@

We assume that medical devices produce 3D images with lateral dimensions that are integer multiples of some powers of two, like 128, 256, 512, etc.
Any cuboidal portion of image is completely determined by the Cartesian indices of its voxels of lowest and highest indices, and extracted by multidimensional array \emph{slicing} as $Image([\ell_x$:$h_x, \ell_y$:$h_y, \ell_z$:$h_z])$.

@@

\paragraph{\emph{Chain coordinates }\label{sec:bbbb}}
We are going to treat each image block independently from each other. Hence we map each image block $\B(i,j,n)$ to the linear \emph{chain} space $C_2$ of dimension $n\times n$, using coordinate vectors $c_{h,k}\in \B^{n\times n} := \{0,1\}^{n\times n}$, where the basis element $c = c_{h,k} \in C_2$ is mapped via Cartesian-to-linear map to the binary vector 
\[
Image(h,k) \mapsto c_{h,k} := [0 \cdots 0\ 1\ 0 \cdots 0] \in \B^{n\times n}
\]
for each $0\leq h,k \leq n$, and where the (single) unit element is in position $nk + h \leq n\times n$.

@@

\paragraph{\emph{Boundary operator }\label{sec:bbbb}}For a fixed block size $n$, the boundary operator $\partial_d : C_d\to C_{d-1}$, with $d\in\{2,3\}$, will be constructed once and for all using the algorithm given in~\cite{}, and inlined in the generated boundary extraction code.

@@

\paragraph{\emph{Sparsity and size of boundary matrix }\label{sec:bbbb}}

@@

Here we refer directly to the 3D case.
Let us call \emph{segment} the bulk content $S$ of interest within the input 3D image of size $(u,v,w)$. We aim to compute the segment boundary $\partial_3 S$. 
First we set the size $n$ of the block, in order to decompose the input $Image(u,v,w)$ into a fair number 
\[
M = \ceil{u/n} \times \ceil{v/n} \times \ceil{w/n} \simeq \frac{uvw}{n^3}
\] of blocks. 
Then we consider each image portion $c_{i,j,k} = S\cap \B(i,j,k,n)$ and compute its (binary) coordinate representation  $[c]_{i,j,k}\in C_3(n,n,n)$. This one is a sparse binary vector of length $n^3$. Then assemble the $M$ representations $c$ of segment portions into a sparse binary matrix $\T{S}$, of dimension $n^d \times M$. Finally compute a matrix $\T{\B}$ of boundary portions of $S$, represented by columns as chain coordinate vectors in $C_2$:
\[
\T{\B} = [\partial_3(n)]\, \T{S}.
\]
where the boundary matrix has dimension $d\,n\,(1+n)^{d-1} \times n^d$.
Of course, the $\T{\B}$ sparse matrix has the same column number $M$ of $\T{S}$, because each column contains the boundary representation of the corresponding $S\cap Box_{i,j,k}{i,j,k}$, and the number of rows of the operator, equal to the dimension of the linear space $C_2$.

\paragraph{Embedding}
A final computational step is needed, in order to embed the 2-chains in Euclidean space $\E^3$ and to assemble the whole resulting surface. In particular, we need to compute the \emph{embedding function} $\mu : C_0 \to \E^3$, where $C_0$ is the space of 0-chains, one-to-one corresponding to the vertices of the extracted surface. The simplest solution is to associate  four 0-cells to each 2-cell of the extracted surface, i.e.~to each non-zero entry in every column of $\T{\B}$.  The $\mu$ function  can be computed by identifying, via  element position in the column, a triple of integer values $0\leq x\leq u$, $0\leq y\leq v$, and $0\leq z\leq w$ for each vertex of the 2-cell.  The mapping can be implemented using a dictionary, that will store the inverse coordinate transformation used at the beginning, i.e.~the one from linear to Cartesian coords, in order of not duplicating the output vertices.

\paragraph{Surface assembling}

All boundary surface subsets $\B_{i,j,k}(S) = \partial_3 S \cap \mbox{Box}_{i,j,k}$, provided by  columns of $\T{S}$, are embedded in the same coordinate space. In formal terms: 
\[
\texttt{Lar}(S) := (\texttt{Geom}(S), \texttt{Top}(S)) = (\texttt{V}, \texttt{CV}),
\]
where, with respect to the \emph{chain complex} $C_3\to C_2\to C_1\to C_0$ induced by the input image $Im$ and segment $S_{i,j,k}$, we get
\begin{align}
\texttt{Geom} &:= \mu(C_0(_{i,j,k})) = \texttt{V},
\\
\texttt{Top} &:= C_3(S) = \T{S} \mapsto \texttt{CV}.
\end{align}
and
\begin{align}
\texttt{Lar}(B_{i,j,k}) &:= (\texttt{Geom}, \texttt{Top}) = (\texttt{W}, \texttt{FW}),
\\
\texttt{Geom} &:= \mu(C_0(B_{i,j,k})) = \texttt{W} \subset \texttt{V},
\\
\texttt{Top} &:= C_2(B_{i,j,k}) = \T{\B}_{i,j,k} \mapsto \texttt{FW} \subset \texttt{FV}.
\end{align}

A translation transformation applied to each vertex subset $\texttt{W}_{i,j,k}$ with translation  vector $\v{t} = [i,j,k]$ will therefore move it in the final space position, so finally giving
\[
\texttt{Lar}(\B) = \oplus_{i,j,k}\texttt{Lar}(\partial_3 S_{i,j,k})) = \oplus_{i,j,k}(\texttt{W}, \texttt{FW}) .
\]

@@

In the computational pipeline introduced in this paper, several steps can be efficiently performed in parallel at image-block level, depending on the embarassingly data parallel nature of the problem. In particular, little effort is needed to separate the problem into a number of parallel tasks $S_{i,j,k}$, using multiarray slicing. The granularity of parallelism, depending on the block size $n$, is further enforced by the computation of a single boundary matrix $[\partial_d(n)]$, in turn depending on $n$, so that the initial communication cost of broadcasting the matrix to nodes can be carefully controlled, and finely tuned depending on the system architecture. The whole approach is appropriate  for SIMD hybrid architectures of CPUs and GPUs, since only the initial block setup of boundary matrix and image slices, as well the final collection of computed surface portions, require inter-process communication.

@@

\paragraph{\emph{Workflow setup}}The functions in this preliminary step include:
\begin{enumerate}

@@

\item depending on the above, best decision for $\emph{size}$ of 3D image blocks $\mathcal{B}$ (bricks); defaults to: $\emph{size}=64$; hence the number of bricks will be $n=\ceil{\ell_1/size}\times\ceil{\ell_2/size}\times\ceil{\ell_3/size}$. 
Hence default value: $n=256$, for images $\mathcal{I} = 512\times 512\times 256$;

@@

\item distribution of matrix $[\partial_B]$, of default size $45$ MB, to all available nodes/cores (workers) using the macro \texttt{@eveywhere}.

@@

\begin{figure}[tbp]
\includegraphics[width=\textwidth]{figs/schema_horizontal.pdf} 
%\includegraphics[scale=1]{input/ircad_comparison.pdf} 
\caption{Workflow of LAR-SURF algorithm}
\label{fig:schema}
\end{figure}

With $size=64$, the number of non-zeros within the sparse matrix $[\partial_{64^3}]$ is $\mathtt{nnz} = 4 792 266$, for a memory size of $9\times \mathtt{nnz}+8\times 262144 \simeq 45$ MB. The memory size of the sparse matrix is computed by considering $8+1$ bytes for non-zero element (which are exactly 6 per row), plus 8 bytes per each index of column start.

\paragraph{\emph{Job enqueuing}}Communication and data synchronization may be managed through \emph{Channels}, which are the FIFO conduits that may provide producer/consumer communication. Overall execution time can be improved if other tasks can be run while a task is being executed, or while waiting for an external service/function to complete. The single work items of this stage follow:
\begin{enumerate}

@@

\paragraph{\emph{3-Chain encoding}}
The interesting part of the \emph{Image} $\mathcal{I}$ is called \emph{Segment} $\mathcal{S}$. The goal of the whole \emph{workflow} is to extract a \emph{boundary model} of $\mathcal{S}$ from $\mathcal{I}$. The portion of $\mathcal{S}$ inside $\mathcal{B}$, will be denoted as $\mathcal{S}(B)$.
Each block $\mathcal{B}$ of the 3D image must by converted into the \emph{coordinate representation} of a vector $\nu\in C_3$ in the linear space  of 3-chains.

@@

\paragraph{\emph{SpMM Multiplication}}
According to the current literature~\cite{} it is more convenient to execute SpMV (sparse matrix-vector) multiplications than SpMSpV (sparse matrix-sparse vector) multiplications. Since we have 256 such jobs (one multiplication per block) to perform in the standard setting of the algorithm\footnote{
Size of block $64^3$; size of image $512^2\times 256$.
}, or more in case of either smaller blocks or image greater than the standard one, this stage must be evidently parallelized and carefully tuned, possibly by using the GPU, if available.
\begin{enumerate}

@@

\paragraph{\emph{2-Chain decoding}}
Each multiplication of $[\partial_B] : C_3 \to C_2$, times a 3-chain $\nu\in C_3$, produces a 2-chain  $\sigma\in C_2$, i.e.~the \emph{coordinate representation} of the \emph{boundary vector} $\sigma\in C_2$.  The inverse of the coding algorithm is executed in the present stage.  This process can also be partially superimposed in time with the previous ones, depending on the size of the memory buffers used to feed the CPU cores or the GPUs and get their results. Some elementary steps follow:

@@

\paragraph{\emph{Assembling and artifact filtering}}
The results of the previous stages can be described as a \emph{collection of sets} of \emph{geometric quadrilaterals (quads)}, each one encoded as an array of quadruples of integer indices, pointing to the linear array of grid vertices associated to the image block $\mathcal{B}$.  In other words, \emph{all quads} of \textbf{each job} are now given in the \textbf{same} \emph{local coordinates}.  Besides to put each partial surface $\mathcal{S}(B) = (\texttt{V}_B, \texttt{FV}_\sigma)$ in the global coordinate system of the image, the present stage must eliminate the redundant boundary features possibly generated at the edges of the partial surface $\mathcal{S}(B)$ within each block $\mathcal{B}$ such that $B \cap \mathcal{I} \not= \emptyset$:

@@

\paragraph{\emph{Smoothing}}
The final smoothing of the generated surfaces cannot be performed block-wise, since this would introduce smoothing artifacts at the block boundaries. Anyway, the Taubin smoothing~\cite{} can be performed in parallel, since for each vertex in the final surface (except eventually the ones on the image $\mathcal{I}$ boundaries) it essentially consist in computing a new position as a proper average of its neighborhood vertices, i.e.~by applying a discrete Laplacian operator.  Some appropriate set workers may so be assigne the task of generating iteratively a new position for the vertices they take cure of. In particular, we have:
\begin{enumerate}

\item Job enqueuing, by writing sets of integers (global linear indices of vertices) in array buffers of appropriate type \texttt{Channel};

\item iterated vectorized computation of proper averages of closed vertices;

\item job dequeuing, by recovering finished tasks from a channel and assembling the results into the embedding function $\mathtt{V}: C_0 \to \E^3$ providing an array of type \texttt{Lar.Points} of \texttt{Float64 $\times$ 3}, with vertex coordinates by column.
\end{enumerate}

>>>>

\usepackage{draftwatermark}
\SetWatermarkText{Draft}
\SetWatermarkScale{2.0}
\SetWatermarkColor[rgb]{0.9,0.9,0.9}

@@

Some basic concepts of solid modeling, and in particular the foundational idea of representation scheme, as well as few basic concepts of algebraic topology, are shortly introduced in this Section, including the computation of the matrix of a boundary operator between chain spaces.

@@

It is possible to show that the \textsc{lar} representation scheme is very expressive, i.e.~that it  has a large domain,  including collections of: line segments, quads, triangles, polygons, meshes;  pixels, voxels, volume images; B-reps, enumerative and decompositive representations of solids. 
In this paper we apply \textsc{lar} methods to computation of boundary representations of solid models from segmentation (labeling) of 3D medical images.
To display a triangulation of  boundary faces  in their proper position in space, the information required is contained in the \emph{geometric chain complex} (GCC):
\[
\mu: C_0\to\E^3,\ (\delta_0, \delta_1, \delta_2)
\qquad\equiv\qquad
\texttt{(geom, top) = (W, (EV, FE, CF))}
\]
The GCC allows to transform the (possibly non connected) boundary 2-cycle of surfaces as a standard B-rep~\cite{shapiroSM:202}. 
The geometry \texttt{geom} is given by
the embedding matrix \texttt{W} of vertices (0-cells), and  topology \texttt{top} by the three sparse matrices (\texttt{EV}, \texttt{FE}, \texttt{CF}) of coboundaries ($\delta_1, \delta_2, \delta_3$) of the chain complex describing a   
space arrangement~\cite{paoluzzi2019finite}.
Note that ordered pairs of letters from \texttt{V,E,F,C}, correspond to \emph{\emph{\texttt{V}}ertices$\to$\emph{\texttt{E}}dges$\to$\emph{\texttt{F}}aces$\to$\emph{\texttt{C}}ells} into the 
\emph{\emph{\texttt{C}}olumn$\to$\emph{\texttt{R}}ow} order of matrix maps of operators.

\subsubsection*{Construction of boundary matrix $\partial_d$}

@@

\begin{figure}[tbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.75\linewidth]{figs/boundary_matrix_4x4x4.pdf} 
   \caption{A binary image of the coboundary operator  $\delta_2 = \partial_3^\top : C_2 \to C_3$, built for a small 3D image with shape $(4,4,4)$. Note that the number of rows equates the cardinality $4\times 4\times 4 = 64$ of the voxel set; the number of columns is $d\,n\,(1+n)^{d-1} = 3\times 4\times 25 = 300$. Of course, the number of non-zeros per row (cardinality of the facet set of a single voxel) is six, whereas the number of non-zeros per column is two, but on boundary facets.}
   \label{fig:boundary_matrix_4x4x4}
\end{figure}

@@

Every boundary chain extracted from an image block $\B(i,j,k,n)$ is a \emph{2-cycle}, i.e., a closed 2-chain---in other words, a 2-chain with empty boundary. Such 2-cycles are joined together by removing the double 2-cells (at the boundaries of adjacent bricks) after having suitably shifted their indices to an unique linear representation of the whole image. The resulting raster surface is made by mutually orthogonal raster facets, that must be smoothed in order to get a fair surface. A linear time and space algorithm for this purpose is the Laplacian smoothing, which iteratively  moves each vertex (0-cell) to the centroid of its neighbors. A well known weakness of this simple algorithm is the asymptotic convergence of the whole mesh to a single point, resulting in unfair size reduction even after few iterations.  Conversely, the Taubin smoothing algorithm~\cite{Taubin:1995:SPA:218380.218473,egst.20001029} alternates two Laplacian smoothing steps with \emph{shrink} and \emph{inflate} effects respectively, with the result of delivering pretty invariant sizes and volume of the smoothed mesh. The best results are obtained on meshes which have small variations of edge length and face angles, like for surfaces extracted from 3D raster images, as in our case.

@@

We assume that medical devices produce 3D images with lateral dimensions that are integer multiples of some powers of two, like 128, 256, 512, etc.
Any cuboidal portion of image is completely determined by the Cartesian indices of its voxels of lowest and highest indices, and extracted by multidimensional array \emph{slicing} as $Image([\ell_x : h_x, \ell_y : h_y, \ell_z : h_z])$.

@@

\subsubsection*{Chain coordinates }\label{sec:bbbb}
We are going to treat each image block independently from each other. Hence we map each image block $\B(i,j,k,n)$ to the linear \emph{chain} space $C_2$ of dimension $n\times n\times n$, using coordinate vectors $c\in \mathbf{2}^{n^d} := \{0,1\}^{n^d}$, where the basis element $c \in C_2$ is mapped via Cartesian-to-linear map to the binary vector 
\[
Image(h,k) \mapsto c_{h,k} := [0 \cdots 0\ 1\ 0 \cdots 0] \in \B^{n\times n}
\]
for each $0\leq h,k \leq n$, and where the (single) unit element is in position $nk + h \leq n\times n$.

@@

\subsubsection*{Boundary operator }\label{sec:bbbb}
For a fixed block size $n$, the boundary operator $\partial_d : C_d\to C_{d-1}$, with $d\in\{2,3\}$, will be constructed once and for all using the algorithm given in~\cite{}, and inlined in the generated boundary extraction code.

@@

\subsubsection*{Sparsity and size of boundary matrix }\label{sec:bbbb}

@@

Here we refer directly to the 3D case.
Let us call \emph{segment} the bulk content $S$ of interest within the input 3D image of size $(u,v,w)$. Our aim is to compute the segment boundary $\partial_3 S$. 
First we set the size $n$ of the block, in order to decompose the input $Image(u,v,w)$ into a fair number of blocks
\[
M = \ceil{u/n} \times \ceil{v/n} \times \ceil{w/n} \simeq \frac{uvw}{n^3}.
\] 
Then, we consider each segment portion $c_{i,j,k} = S\cap \B(i,j,k,n)$ and compute its local coordinate representation  $[c]_{i,j,k}\in C_3(n,n,n)$. This one is a sparse binary vector of length $n^3$. Then, assemble the $M$ representations $c$ of segment portions into a sparse binary matrix $\T{S}$, of dimension $n^d \times M$, with $d=3$. Finally, compute a matrix $\T{\textbf{B}}$ of boundary portions of $S$, represented by columns as chain coordinate vectors in $C_2$:
\[
\T{\textbf{B}} = [\partial_3(n)]\, \T{S}.
\]
where the boundary matrix has dimension $n^d \times dn(n+1)^{d-1}$.
Of course, the $\T{\textbf{B}}$ sparse matrix has the same column number $M$ of $\T{S}$, because each column contains the boundary representation of the corresponding $S \cap \B(i,j,k,n)$, and the number of matrix rows is the dimension $n^d$ of the linear space $C_2$ of 2-chains.

\subsubsection*{Embedding}
A final computational step is needed, in order to embed the 2-chains in $\E^3$ space and to assemble the whole resulting surface. In particular, we need to compute the \emph{embedding function} $\mu : C_0 \to \E^3$, where $C_0$ is the space of 0-chains, one-to-one with the vertices of the extracted surface. The simplest solution is to associate  four 0-cells to each 2-cell of the extracted surface, i.e.~to each non-zero entry in every column of $\T{\textbf{B}}$.  The $\mu$ function  can be computed by identifying, via  element position in the column, a triple of integer values $0\leq x\leq u$, $0\leq y\leq v$, and $0\leq z\leq w$ for each vertex of the 2-cell.  The mapping can be implemented using a dictionary, that will store the inverse coordinate transformation used at the beginning, i.e.~the one from linear to Cartesian coords, in order of not duplicating the output vertices.

\subsubsection*{Surface assembling}

All boundary surface subsets $B(i,j,k) = \partial_3 S \cap \B(i,j,k)$, provided by  columns of $\T{B}$, are embedded in the same coordinate space. In formal terms, using the standard terminology of LAR scheme: 
\[
\texttt{Lar}(S) := (\texttt{Geom}(S), \texttt{Top}(S)) = (\texttt{V}, \texttt{CV}),
\]
where, with respect to the \emph{chain complex} $C_3\to C_2\to C_1\to C_0$ induced by the input image $Im$ and segment portion $S_{i,j,k}$, we get
\begin{align}
\texttt{Geom} &:= \mu(C_0) = \texttt{V},
\\
\texttt{Top} &:= C_3(S) = \T{S} \mapsto \texttt{CV}.
\end{align}
and
\begin{align}
\texttt{Lar}(B_{i,j,k}) &:= (\texttt{Geom}(B_{i,j,k}), \texttt{Top}(B_{i,j,k})) = (\texttt{W}, \texttt{FW}),
\\
\texttt{Geom} &:= \mu(C_0(B_{i,j,k})) = \texttt{W} \subset \texttt{V},
\\
\texttt{Top} &:= C_2(B_{i,j,k}) = \T{\textbf{B}}_{i,j,k} \mapsto \texttt{FW} \subset \texttt{FV}.
\end{align}

A translation transformation applied to each vertex subset $\texttt{W}_{i,j,k}$ with translation  vector $\v{t} = [i,j,k]$ will therefore move it in the final space position, so finally giving
\[
\texttt{Lar}(\textbf{B}) = \oplus_{i,j,k}\texttt{Lar}(\partial_3 S_{i,j,k})) = \oplus_{i,j,k}(\texttt{W}, \texttt{FW}) .
\]

@@

In the computational pipeline introduced in this paper, several steps can be efficiently performed in parallel at image-block level, depending on the embarassingly data parallel nature of the problem. In particular, little effort is needed to separate the problem into a number of parallel tasks $S_{i,j,k}$, using multiarray slicing. The granularity of parallelism, depending on the block size $n$, is further enforced by the computation of a single boundary matrix $[\partial_d(n)]$ depending on $n$, so that the initial communication cost of broadcasting the matrix to nodes can be carefully controlled, and finely tuned depending on the system architecture. The whole approach is appropriate  for SIMD hybrid architectures of CPUs and GPUs, since only the initial block setup of boundary matrix and image slices, as well the final collection of computed surface portions, require inter-process communication.

@@

\subsubsection*{Workflow setup}\label{}The functions in this preliminary step include:
\begin{enumerate}

@@

\item depending on the above, best decision for $\emph{size}$ of 3D image blocks $\mathcal{B}$ (bricks). With defaults $\emph{size}=64$,  the number of bricks will be $n=\ceil{\ell_1/size}\times\ceil{\ell_2/size}\times\ceil{\ell_3/size}$. 
Hence the default number of bricks is $n=256$, for standard images $512\times 512\times 256$;

@@

\item distribution of matrix $[\partial_B]$, of default size $45$ MB, to all available nodes/cores (workers) using the Julia macro \texttt{@eveywhere}.

@@

\begin{figure}[tbp]
\includegraphics[width=\textwidth]{figs/schema_horizontal.pdf} 
%\includegraphics[scale=1]{input/ircad_comparison.pdf} 
\caption{Workflow of \textsc{lar-surf} algorithm}
\label{fig:schema}
\end{figure}

With $size=64$, the number of non-zeros within the sparse matrix $[\partial(64^3)]$ is $\mathtt{nnz} = 4 792 266$, for a memory size of $9\times \mathtt{nnz}+8\times 262144 \simeq 45$ MB. The memory size of the sparse matrix is computed by considering $8+1$ bytes for non-zero element (which are exactly 6 per row), plus 8 bytes per each index of column start.

\subsubsection*{Job enqueuing}\label{}Communication and data synchronization may be managed through \emph{Channels}, which are the FIFO conduits that may provide producer/consumer communication. Overall execution time can be improved if other tasks can be run while a task is being executed, or while waiting for an external service/function to complete. The single work items of this stage follow:
\begin{enumerate}

@@

\subsubsection*{3-Chain encoding}\label{}
The interesting part of the \emph{Image} $\mathcal{I}$ is called \emph{Segment} $\mathcal{S}$. The goal of the whole \emph{workflow} is to extract a \emph{boundary model} of $\mathcal{S}$ from $\mathcal{I}$. The portion of $\mathcal{S}$ inside $\mathcal{B}$, will be denoted as $\mathcal{S}(B)$.
Each block $\mathcal{B}$ of the 3D image must by converted into the \emph{coordinate representation} of a vector $\nu\in C_3$ in the linear space  of 3-chains.

@@

\subsubsection*{SpMM Multiplication}\label{}
According to the current literature~\cite{} it is more convenient to execute SpMV (sparse matrix-vector) multiplications than SpMSpV (sparse matrix-sparse vector) multiplications. Since we have 256 such jobs (one multiplication per block) to perform in the standard setting of the algorithm\footnote{
Size of block $64^3$; size of image $512^2\times 256$.
}, or more in case of either smaller blocks or image greater than the standard one, this stage must be evidently parallelized and carefully tuned, possibly by using the GPU, if available.
\begin{enumerate}

@@

\subsubsection*{2-Chain decoding}\label{}
Each multiplication of $[\partial_B] : C_3 \to C_2$, times a 3-chain $\nu\in C_3$, produces a 2-chain  $\sigma\in C_2$, i.e.~the \emph{coordinate representation} of the \emph{boundary vector} $\sigma\in C_2$.  The inverse of the coding algorithm is executed in the present stage.  This process can also be partially superimposed in time with the previous ones, depending on the size of the memory buffers used to feed the CPU cores or the GPUs and get their results. Some elementary steps follow:

@@

\subsubsection*{Assembling and artifact filtering}\label{}
The results of the previous stages can be described as a \emph{collection of sets} of \emph{geometric quadrilaterals (quads)}, each one encoded as an array of quadruples of integer indices, pointing to the linear array of grid vertices associated to the image block $\mathcal{B}$.  In other words, \emph{all quads} of \textbf{each job} are now given in the \textbf{same} \emph{local coordinates}.  Besides to put each partial surface $\mathcal{S}(B) = (\texttt{V}_B, \texttt{FV}_\sigma)$ in the global coordinate system of the image, the present stage must eliminate the redundant boundary features possibly generated at the edges of the partial surface $\mathcal{S}(B)$ within each block $\mathcal{B}$ such that $B \cap \mathcal{I} \not= \emptyset$:

@@

\subsubsection*{Smoothing}\label{}
The final smoothing of the generated surfaces cannot be performed block-wise, since this would introduce smoothing artifacts at the block boundaries. Anyway, the Taubin smoothing~\cite{} can be performed in parallel, since for each vertex in the final surface (except eventually the ones on the image $\mathcal{I}$ boundaries) it essentially consist in computing a new position as a proper average of its neighborhood vertices, i.e.~by applying a discrete Laplacian operator.  Some appropriate set of workers may so be assigned the task of generating iteratively a new position for the vertices they take cure of. In particular, we have:
\begin{enumerate}

\item Job enqueuing, by writing sets of integers (global linear indices of vertices) in array buffers of type \texttt{Channel};

\item vectorized computation of proper averages of near vertices;

\item job dequeuing, by recovering finished tasks from a channel and assembling the results into the embedding function $\mathtt{V}: C_0 \to \E^3$, providing an array of type \texttt{Lar.Points} of \texttt{Float64 $\times$ 3}, with vertex coordinates by column.
\end{enumerate}
